---
title: "MADA Project_UOWN"
author: "C. Coleman"
date: "9/16/2021"
output: html_document
---

# *Please run the "MADA_Project" file in this folder to ensure use of the correct working directory.*

# Overview: Comparison of Stream Health indicators

### The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN). This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed. Trained freshwater scientists assist in training members of the community to collect both chemical and biological stream characteristics that are useful in describing stream health. All data is compliled and posted in raw form on the UOWN website. Of interest to me are the following variables: WSID (ID number),	biological_score,	conductivity, turbidity,	po4.mgL,	no3.mgL,	pH,	temperature,	e.coli.cfu,	month,	year,	day. The raw data shows 3,075 observations before cleaning. However, while cleaning I had to eliminate "po4.mgl" and "temperature" du to significant amounts of missing data.

### For quick reference, here is hoe to load the raw data:

#### Load import packages:
```{r}
library(tidyverse)
library(readxl)
```

#### Lodaing data and providing a summary for the raw data
```{r}
data_location <- here::here("data","raw_data","UOWN_data_master_spring2021 (2).xlsx")

rawdata <- readxl::read_excel(data_location)

glimpse(data_cleaning)
```

# Getting Started From Scratch

This file contains four main sub-folders titled "code", "data", "products", and "results". The "code" folder contains all the code used to process, explore and analyze raw data from the UOWN Site. The "data" folder contains the raw and processed data for this project. The processed data was produced using the processing script located in the "data" folder. The "results" folder contains all figures and tables created from the data exploration and analysis code, located in the "code" folder. Finally, the "product" folder contains the final manuscript file, supplementary file, and all accompanying files used to create a reference library in Bibtex and format the manuscript word document.

# Workflow:

1. Once in the R project file located in this file, navigate into the "code" file. Here, run the processing script to create the processed data. This will now be located in the "data" folder, in the sub folder "processed_data".

2. In the "code" file, run the "data_exploration" file to run the introductory linear relationships and create figures showing variable relationships. The .RMD file will save figures in the "results folder".

3. Next, run the "analysis_uown" file to run the outlier and MIDO LASSO linear analysis models. Results from this file will be saved in the "results" file.

4. Finally, run the "Manuscript" and "Supplementary_Data" files to produce the final article and supplementary report for the UOWN data analysis project.

# Question:

When monitoring freshwater streams, stream health is typically measured by its physical characteristics (chemistry, surrounding land use area, channelization, etc.); Microbial (fecal indicator), and ecology (diversity of indicator macroinvertebrates). However, these categories are usually analyzed separately to assess stream health. As such, my question is how well are these parameters correlated with each other. Specifically, can you use one of the indicators mentioned (physical, microbial, and ecological) to describe both of the others to make a statement about overall stream health?

I would expect that because the all three are correlated with urbanization/anthropocentric impact, a negative impact in one indicator category will indicate a correlated negative impact in the other categories. The only caveat is that because level of noise in stream health data, significant correlations may be difficult to find.


