---
title: "Code_Dump_Project"
author: "C. Coleman"
date: "10/28/2021"
output: html_document
---
_Additionally, we will eventually be removing outliers from a series of relationships between many predictors and one outcome. You can imagine that because we are looking at many predictors with one outcome, we will either have some overlap in the outcome values we are trying to remove or potentially remove every single outcome value if we try to do this in the same data set. As such, we will need to split the MIDO, NORO, and BICO dataframes into individual dataframes that contain our outcome of interest and each single variable we are trying to assess a relationship. We will do this below._
```{r, MIDO}
bs_con_df <- MIDO %>%
  select("biological_score", "conductivity.uscm")
bs_ecfu_df <- MIDO %>%
  select("biological_score", "e.coli.cfu")
bs_no3_df <- MIDO %>%
  select("biological_score", "no3.mgL")
bs_turb_df <- MIDO %>%
  select("biological_score", "turbidity.ntu")
ecfu_con_df <- MIDO %>%
  select("e.coli.cfu", "conductivity.uscm")
ecfu_no3_df <- MIDO %>%
  select("e.coli.cfu", "no3.mgL")
ecfu_turb_df <- MIDO %>%
  select("e.coli.cfu", "turbidity.ntu")
```
</br>
```{r, NORO}
n_bs_con_df <- NORO %>%
  select("biological_score", "conductivity.uscm")
n_bs_ecfu_df <- NORO %>%
  select("biological_score", "e.coli.cfu")
n_bs_no3_df <- NORO %>%
  select("biological_score", "no3.mgL")
n_bs_turb_df <- NORO %>%
  select("biological_score", "turbidity.ntu")
n_ecfu_con_df <- NORO %>%
  select("e.coli.cfu", "conductivity.uscm")
n_ecfu_no3_df <- NORO %>%
  select("e.coli.cfu", "no3.mgL")
n_ecfu_turb_df <- NORO %>%
  select("e.coli.cfu", "turbidity.ntu")
```
</br>
```{r, BICO}
b_bs_con_df <- BICO %>%
  select("biological_score", "conductivity.uscm")
b_bs_ecfu_df <- BICO %>%
  select("biological_score", "e.coli.cfu")
b_bs_no3_df <- BICO %>%
  select("biological_score", "no3.mgL")
b_bs_turb_df <- BICO %>%
  select("biological_score", "turbidity.ntu")
b_ecfu_con_df <- BICO %>%
  select("e.coli.cfu", "conductivity.uscm")
b_ecfu_no3_df <- BICO %>%
  select("e.coli.cfu", "no3.mgL")
b_ecfu_turb_df <- BICO %>%
  select("e.coli.cfu", "turbidity.ntu")
```


</br>

# Remove Outliers
_To remove outliers, we must first create a new vector we will call "outliers_relationship", then we will just subtract it out of the original data frames (MIDO, NORO, BICO)_
```{r, MIDO}
Outliers_MIDO_bs_con <- as.numeric(names(bs_con_cd)[(bs_con_cd > (4/sample_size_MIDO))])
Outliers_MIDO_bs_ecfu <- as.numeric(names(bs_ecfu_cd)[(bs_ecfu_cd > (4/sample_size_MIDO))])
Outliers_MIDO_bs_no3 <- as.numeric(names(bs_no3_cd)[(bs_no3_cd > (4/sample_size_MIDO))])
Outliers_MIDO_bs_turb <- as.numeric(names(bs_turb_cd)[(bs_turb_cd > (4/sample_size_MIDO))])
Outliers_MIDO_ecfu_con <- as.numeric(names(ecfu_con_cd)[(ecfu_con_cd > (4/sample_size_MIDO))])
Outliers_MIDO_ecfu_no3 <- as.numeric(names(ecfu_no3_cd)[(ecfu_no3_cd > (4/sample_size_MIDO))])
Outliers_MIDO_ecfu_turb <- as.numeric(names(ecfu_turb_cd)[(ecfu_turb_cd > (4/sample_size_MIDO))])
```
</br>
```{r, NORO}
Outliers_NORO_bs_con <- as.numeric(names(n_bs_con_cd)[(n_bs_con_cd > (4/sample_size_NORO))])
Outliers_NORO_bs_ecfu <- as.numeric(names(n_bs_ecfu_cd)[(n_bs_ecfu_cd > (4/sample_size_NORO))])
Outliers_NORO_bs_no3 <- as.numeric(names(n_bs_no3_cd)[(n_bs_no3_cd > (4/sample_size_NORO))])
Outliers_NORO_bs_turb <- as.numeric(names(n_bs_turb_cd)[(n_bs_turb_cd > (4/sample_size_NORO))])
Outliers_NORO_ecfu_con <- as.numeric(names(n_ecfu_con_cd)[(n_ecfu_con_cd > (4/sample_size_NORO))])
Outliers_NORO_ecfu_no3 <- as.numeric(names(n_ecfu_no3_cd)[(n_ecfu_no3_cd > (4/sample_size_NORO))])
Outliers_NORO_ecfu_turb <- as.numeric(names(n_ecfu_turb_cd)[(n_ecfu_turb_cd > (4/sample_size_NORO))])
```
</br>
```{r, BICO}
Outliers_BICO_bs_con <- as.numeric(names(b_bs_con_cd)[(b_bs_con_cd > (4/sample_size_BICO))])
Outliers_BICO_bs_ecfu <- as.numeric(names(b_bs_ecfu_cd)[(b_bs_ecfu_cd > (4/sample_size_BICO))])
Outliers_BICO_bs_no3 <- as.numeric(names(b_bs_no3_cd)[(b_bs_no3_cd > (4/sample_size_BICO))])
Outliers_BICO_bs_turb <- as.numeric(names(b_bs_turb_cd)[(b_bs_turb_cd > (4/sample_size_BICO))])
Outliers_BICO_ecfu_con <- as.numeric(names(b_ecfu_con_cd)[(b_ecfu_con_cd > (4/sample_size_BICO))])
Outliers_BICO_ecfu_no3 <- as.numeric(names(b_ecfu_no3_cd)[(b_ecfu_no3_cd > (4/sample_size_BICO))])
Outliers_BICO_ecfu_turb <- as.numeric(names(b_ecfu_turb_cd)[(b_ecfu_turb_cd > (4/sample_size_BICO))])
```
</br>
_With the outlier vectors created, all we need to do now is subtract them from the original cleaned data frames (MIDO, NORO, BICO)._
```{r, MIDO}
bs_con_df <- bs_con_df
bs_ecfu_df <- bs_ecfu_df
bs_no3_df <- bs_no3_df
bs_turb_df <- bs_turb_df
ecfu_con_df <- ecfu_con_df
ecfu_no3_df <- ecfu_no3_df
ecfu_turb_df <- ecfu_turb_df
```
</br>
```{r, NORO}
n_bs_con_df <- 
n_bs_ecfu_df <- 
n_bs_no3_df <- 
n_bs_turb_df <- 
n_ecfu_con_df <- 
n_ecfu_no3_df <- 
n_ecfu_turb_df <- 
```
</br>
```{r, BICO}
b_bs_con_df <- 
b_bs_ecfu_df <- 
b_bs_no3_df <- 
b_bs_turb_df <- 
b_ecfu_con_df <- 
b_ecfu_no3_df <- 
b_ecfu_turb_df <- 
```
```{r, MIDO}
MIDO2 <- MIDO %>%
  [-Outliers_MIDO_bs_con] 
  MIDO[-Outliers_MIDO_bs_ecfu, ] 
  MIDO[-Outliers_MIDO_bs_no3, ] 
  MIDO[-Outliers_MIDO_bs_turb, ] 
  MIDO[-Outliers_MIDO_ecfu_con, ] 
  MIDO[-Outliers_MIDO_ecfu_no3, ] 
  MIDO[-Outliers_MIDO_ecfu_turb, ]
```

</br>
# Linear Regression Analysis Corrected for Outliers
</br>
## MIDO
</br>
#### Biological Indicator:
_Biological Vs Conductivity_
```{r}
BS_Con <- MIDO %>% 
  ggplot(aes(x=(conductivity.uscm), 
             y=(biological_score))) +
  geom_point(color = "red") +
  geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

print(BS_Con)
```

#### Now run linear regression for BS_Con.
```{r}
BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=MIDO)

summary(BS_Con_lm)
```

#### With *m = -0.017* and *p = 0.49*, there is a *negative*, *non-significant* correlation between *biological score* and *conductivity*.

</br>

#### Biological Vs Turbidity
```{r}
BS_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(BS_Turb)
```

#### Now run linear regression for BS_Turb.
```{r}
BS_Turb_lm <- lm(biological_score ~ turbidity.ntu, data=MIDO)

summary(BS_Turb_lm)
```
#### With *m = 0.11* and *p = 0.13*, there is a *positive*, *non-significant* correlation between *biological score* and *turbidity*.

</br>

#### Biological Vs Nitrate Concentration
```{r}
BS_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(BS_no3)
```

#### Now run linear regression for BS_no3.
```{r}
BS_no3_lm <- lm(biological_score ~ no3.mgL, data=MIDO)

summary(BS_no3_lm)
```

#### With *m = 0.11* and *p = 0.27*, there is a *positive*, *non-significant* correlation between *biological score* and *no3*.

</br>

#### Biological Vs E. Coli CFU
```{r}
BS_ECFU <- MIDO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(BS_ECFU)
```

#### Now run linear regression for BS_ECFU.
```{r}
BS_ECFU_lm <- lm(biological_score ~ e.coli.cfu, data=MIDO)

summary(BS_ECFU_lm)
```
#### With *m = -0.0007* and *p = 0.39*, there really isn't a relationship at all between *biological score* and *ECFU*.

</br>

#### Now let's use the gridExtra package to view all the MIDO Biological Indicator outcome relationship graphs together.
```{r}
grid.arrange(BS_Con, BS_ECFU, BS_no3, BS_Turb, ncol = 2)
```


### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
ECFU_Con <- MIDO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(ECFU_Con)
```

#### Now run linear regression for ECFU_Con.
```{r}
ECFU_Con_lm <- lm(e.coli.cfu ~ conductivity.uscm, data=MIDO)

summary(ECFU_Con_lm)
```
#### With *m = -3.635* and *p = 0.28*, there is a *negative*, *non-significant* correlation between *ECFU* and *Conductivity*.

</br>

#### E. Coli Vs Turbidity
```{r}
ECFU_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(ECFU_Turb)
```

#### Now run linear regression for ECFU_Turb.
```{r}
ECFU_Turb_lm <- lm(e.coli.cfu ~ turbidity.ntu, data=MIDO)

summary(ECFU_Turb_lm)
```
#### With *m = 18.6* and *p = 0.07*, there is a *positive*, *non-significant* correlation between *ECFU* and *turbidity*.

</br>

#### E. Coli Vs Nitrate Concentration
```{r}
ECFU_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(ECFU_no3)
```

#### Now run linear regression for ECFU_no3.
```{r}
ECFU_no3_lm <- lm(e.coli.cfu ~ no3.mgL, data=MIDO)

summary(ECFU_no3_lm)
```

#### With *m = -10.61* and *p = 0.60*, there is a *negative*, *non-significant* correlation between *ECFU* and *no3*.

</br>

#### Now let's use the gridExtra package to view all the MIDO E. coli CFU outcome relationship graphs together.
```{r}
grid.arrange(ECFU_Con, ECFU_no3, ECFU_Turb, ncol = 2)
```
</br>

## NORO

</br>

### *Biological Indicator:*

#### Biological Vs Conductivity
```{r}
N_BS_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(N_BS_Con)
```

#### Now run linear regression for N_BS_Con.
```{r}
N_BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=NORO)

summary(N_BS_Con_lm)
```

#### With *m = -0.023* and *p = 0.0004*, there is a *negative*, *significant* correlation between *biological score* and *conductivity*.

</br>

#### Biological Vs Turbidity
```{r}
N_BS_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(N_BS_Turb)
```

#### Now run linear regression for N_BS_Turb.
```{r}
N_BS_Turb_lm <- lm(biological_score ~ turbidity.ntu, data=NORO)

summary(N_BS_Turb_lm)
```

#### With *m = 0.17* and *p = 0.45*, there is a *positive*, *non-significant* correlation between *bilogical score* and *turbidity*.

</br>

#### Biological Vs Nitrate Concentration
```{r}
N_BS_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(N_BS_no3)
```

#### Now run linear regression for N_BS_no3.
```{r}
N_BS_no3_lm <- lm(biological_score ~ no3.mgL, data=NORO)

summary(N_BS_no3_lm)
```

#### With *m = -0.56* and *p = 0.08*, there is a *negative*, *non-significant* correlation between *biological score* and *no3*.

</br>

#### Now let's use the gridExtra package to view all the NORO Biological Indicator outcome relationship graphs together.
```{r}
grid.arrange(N_BS_Con, N_BS_ECFU, N_BS_no3, N_BS_Turb, ncol = 2)
```
</br>

#### Biological Vs E. Coli CFU
```{r}
N_BS_ECFU <- NORO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(N_BS_ECFU)
```

#### Now run linear regression for N_BS_ECFU.
```{r}
N_BS_ECFU_lm <- lm(biological_score ~ e.coli.cfu, data=NORO)

summary(N_BS_ECFU_lm)
```

#### With *m = 7.274e-05* and *p = 0.69*, there is not relationship between *Bbiological score* and *ECFU*.

</br>

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
N_ECFU_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(N_ECFU_Con)
```

#### Now run linear regression for N_ECFU_Con.
```{r}
N_ECFU_Con_lm <- lm(e.coli.cfu ~ conductivity.uscm, data=NORO)

summary(N_ECFU_Con_lm)
```

#### With *m = -8.632* and *p = 0.20*, there is a *negative*, *non-significant* correlation between *ECFU* and *conductivity*.

</br>

#### E. Coli Vs Turbidity
```{r}
N_ECFU_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black")


print(N_ECFU_Turb)
```

#### Now run linear regression for N_ECFU_Turb.
```{r}
N_ECFU_Turb_lm <- lm(e.coli.cfu ~ turbidity.ntu, data=NORO)

summary(N_ECFU_Turb_lm)
```

#### With *m = 490* and *p = 0.02*, there is a *positive*, *significant* correlation between *ECFU* and *Turbidity*.

</br>

#### E. Coli Vs Nitrate Concentration
```{r}
N_ECFU_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(N_ECFU_no3)
```

#### Now run linear regression for N_ECFU_no3.
```{r}
N_ECFU_no3_lm <- lm(e.coli.cfu ~ no3.mgL, data=NORO)

summary(N_ECFU_no3_lm)
```

#### With *m = -213.1* and *p = 0.49*, there is a *negative*, *non-significant* correlation between *ECFU* and *no3*.

</br>

#### Now let's use the gridExtra package to view all the NORO E. coli CFU outcome relationship graphs together.
```{r}
grid.arrange(N_ECFU_Con, N_ECFU_no3, N_ECFU_Turb, ncol = 2)
```

</br>

## BICO

</br>

### *Biological Indicator:*

#### Biological Vs Conductivity
```{r}
B_BS_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(B_BS_Con)
```

#### Now run linear regression for B_BS_Con.
```{r}
B_BS_Con_lm <- lm(biological_score ~ conductivity.uscm, data=BICO)

summary(B_BS_Con_lm)
```

#### With *m = 0.0046* and *p = 0.86*, there is a *positive*, *non-significant* correlation between *biological score* and *conductivity*.

</br>

#### Biological Vs Turbidity
```{r}
B_BS_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(B_BS_Turb)
```

#### Now run linear regression for B_BS_Turb.
```{r}
B_BS_Turb_lm <- lm(biological_score ~ turbidity.ntu, data=BICO)

summary(B_BS_Turb_lm)
```

#### With *m = -0.13* and *p = 0.44*, there is a *negative*, *non-significant* correlation between *biological score* and *turbidity*.

</br>

#### Biological Vs Nitrate Concentration
```{r}
B_BS_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(B_BS_no3)
```

#### Now run linear regression for B_BS_no3.
```{r}
B_BS_no3_lm <- lm(biological_score ~ no3.mgL, data=BICO)

summary(B_BS_no3_lm)
```

#### With *m = -0.30* and *p = 0.84*, there is a *negative*, *non-significant* correlation between *biological score* and *no3*.

</br>

#### Biological Vs E. Coli CFU
```{r}
B_BS_ECFU <- BICO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(B_BS_ECFU)
```

#### Now run linear regression for B_BS_ECFU.
```{r}
B_BS_ECFU_lm <- lm(biological_score ~ e.coli.cfu, data=BICO)

summary(B_BS_ECFU_lm)
```

#### With *m = -0.00041* and *p = 0.19*, there is no relationship between *biological score* and *ECFU*.

</br>

#### Now let's use the gridExtra package to view all the BICO Biological Indicator outcome relationship graphs together.
```{r}
grid.arrange(B_BS_Con, B_BS_ECFU, B_BS_no3, B_BS_Turb, ncol = 2)
```

</br>

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
B_ECFU_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(B_ECFU_Con)
```

#### Now run linear regression for B_ECFU_Con.
```{r}
B_ECFU_Con_lm <- lm(e.coli.cfu ~ conductivity.uscm, data=BICO)

summary(B_ECFU_Con_lm)
```

#### With *m = 2.28* and *p = 0.90*, there is a *positive*, *non-significant* correlation between *ECFU* and *conductivity*.

</br>

#### E. Coli Vs Turbidity
```{r}
B_ECFU_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(B_ECFU_Turb)
```

#### Now run linear regression for B_ECFU_Turb.
```{r}
B_ECFU_Turb_lm <- lm(e.coli.cfu ~ turbidity.ntu, data=BICO)

summary(B_ECFU_Turb_lm)
```

#### With *m = 261.5* and *p = 0.01*, there is a *positive*, *significant* correlation between *ECFU* and *turbidity*.

</br>

#### E. Coli Vs Nitrate Concentration
```{r}
B_ECFU_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')


print(B_ECFU_no3)
```

#### Now run linear regression for B_ECFU_no3.
```{r}
B_ECFU_no3_lm <- lm(e.coli.cfu ~ no3.mgL, data=BICO)

summary(B_ECFU_no3_lm)
```

#### With *m = -983.6* and *p = 0.33*, there is a *negative*, *non-significant* correlation between *ECFU* and *no3*.

</br>

#### Now let's use the gridExtra package to view all the BICO E. coli CFU outcome relationship graphs together.
```{r}
grid.arrange(B_ECFU_Con, B_ECFU_no3, B_ECFU_Turb, ncol = 2)
```
</br>
